---
title: "hw2_jo2570"
author: "Jai Oberoi"
date: "02/10/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Running the tidyverse package 

```{r}
library(tidyverse)
```

Problem 1: 

Reading in the Mr. Trash Wheel dataset:
Cleaning the names 
Deleting last 3 columns (15, 16 and 17)
Dropping rows that detail mothly total (not specific to a dumpster)
Converting sports balls variable to an integer value 

```{r}
mr_trash_wheel = 
  readxl::read_excel(path = "./data/trash_wheel_data.xlsx") %>%
  janitor::clean_names() %>% 
  select(-x15, -x16, -x17) %>%
  drop_na(dumpster) %>% 
  mutate(sports_balls = as.integer(sports_balls))
```

Reading in the precipitation data for 2017:
Adding a variable year = 2017 
Deleting rows without precipiatation data 

```{r}
ppt_2017 = 
  readxl::read_excel("./data/trash_wheel_data.xlsx", sheet = 6, range = "A2:B14") %>% 
  janitor::clean_names() %>%
  mutate(year = 2017)
```

Reading in the precipitation data for 2017:
Adding a variable year = 2017 
Deleting rows without precipiatation data 

```{r}
ppt_2018 = 
  readxl::read_excel("./data/trash_wheel_data.xlsx", sheet = 5, range = "A2:B14") %>% 
  janitor::clean_names() %>%
  mutate(year = 2018)
```

Merging the datasets and convering month into a character variable 

```{r}
ppt_merged =
  bind_rows(ppt_2017, ppt_2018) %>%
  mutate(month = month.abb[month])
```

the size of the dataframe `mr_trash_wheel` is `r nrow(mr_trash_wheel)` rows (observations) and `r ncol(mr_trash_wheel)` columns (variables). we see that the data shows us how mych trash (weight and volume) a dumpster processes on a partcular date. we also have the number of diff items processed as variables and the number of houses powered. 
the median number of sports balls for 2017 `r median(pull(filter(mr_trash_wheel,year == 2017),sports_balls))`

the size of the dataframe `ppt_merged` is `r nrow(ppt_merged)` rows (observations) and `r ncol(ppt_merged)` columns (variables).
the total precipiation for 2018 is `r sum(pull(ppt_2018, total))`

-- 

Problem 2: 

Reading in and cleaning the pols-month data:
Seperating mon variable into year, month, day 
Creating a variable called president which tells indicates whether the president was a republican ("gop") or a democrat ("dem") 
Dropping/Deleting day, prez_gop and prez_dem

```{r}
pols_month_data = 
  read_csv(file = "./data/fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day")) %>%
  mutate(pesident = ifelse(prez_gop, "gop", "dem"),
         month = as.numeric(month),
         month = month.abb[month],
         year = as.numeric(year)) %>%
  select(-day, -prez_gop, -prez_dem)
```

Reading in and cleaning the snp data: 
Seperating mon variable into year, month, day 
Converting year and month into numeric variables 
Dropping the day column 
Arranging data by year, month 

```{r}
snp_data = 
  read_csv(file = "./data/fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date, into = c("month", "day", "year")) %>%
  mutate(year = as.numeric(year),
         month = as.numeric(month),
         month = month.abb[month]) %>%
  select(year, month, -day, close) %>%
  arrange(year, month)
```

Reading in and cleaning the unemployment data: 
Reducing the number of columns by creating a month column and a rate column 

```{r}
umempl_data = 
  read_csv(file = "./data/fivethirtyeight_datasets/unemployment.csv") %>%
  pivot_longer((Jan:Dec),
               names_to = "month",
               values_to =  "rate") %>%
  rename(year = Year)
```

Merging the snp dataset to the pols dataset 

```{r}
snp_pols_data = 
  left_join(pols_month_data, snp_data, by = c("year", "month"))

```

Merging the unempl dataset to the merged dataset snp_pols: 

```{r}
full_data = 
    left_join(snp_pols_data, umempl_data, by = c("year", "month"))
```

the size of the dataframe `pols_month_data` is `r nrow(pols_month_data)` rows and `r ncol(pols_month_data)` columns. 

the size of the dataframe `snp_data` is `r nrow(snp_data)` rows and `r ncol(snp_data)` columns.  

the size of the dataframe `umempl_data` is `r nrow(umempl_data)` rows and `r ncol(umempl_data)` columns.  

the size of the dataframe `snp_pols_data` is `r nrow(snp_pols_data)` rows and `r ncol(snp_pols_data)` columns.  

the size of the dataframe `full_data` is `r nrow(full_data)` rows and `r ncol(full_data)` columns. the full dataset is a result of presidential data, snp data and umployment data merged by the year and month variables. the range of years is `r range(pull(full_data, year)) ` and some key variable are president, which tells us whether the resident was dem or gop and the unpmoyment rate. 

-- 

Problem 3: 

Reading in and cleaning pop_baby_names data 

```{r}
pop_baby_names_data = 
  read_csv(file = "./data/popular_baby_names.csv") %>%
  janitor::clean_names() %>%
  distinct() %>%
    mutate(gender = str_to_lower(gender),
           childs_first_name = str_to_lower(childs_first_name),
           ethnicity = recode(ethnicity, "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER", "BLACK NON HISP" = "BLACK NON HISPANIC", "WHITE NON HISP" = "WHITE NON HISPANIC"),
           ethnicity = str_to_lower(ethnicity))
```

To get a well-structured, reader-friendly table showing the rank in popularity of the name “Olivia” as a female baby name over time

```{r}
pop_baby_names_data %>%
  filter(childs_first_name == "olivia") %>%
  select(-count) %>%
  pivot_wider(names_from = year_of_birth, 
              values_from = rank) %>% 
  knitr::kable()
```

To get a table showing the most popular name among male children over time: 

```{r}
pop_baby_names_data %>%
  filter(gender == "male", rank == 1) %>%
  select(-count) %>%
  pivot_wider(names_from = year_of_birth, 
              values_from = childs_first_name) %>% 
  knitr::kable()
```

To get a scatter plot of the number of children with a name against the rank in popularity of that name: 

```{r}
scatter_data = pop_baby_names_data %>%
  filter(gender == "male", ethnicity == "white non hispanic", year_of_birth == 2016)

ggplot(scatter_data, aes(x = rank, y = count)) +
geom_point()
```
