---
title: "hw2_jo2570"
author: "Jai Oberoi"
date: "02/10/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Running the tidyverse package 
```{r}
library(tidyverse)
```

Problem 1: 

Reading in the Mr. Trash Wheel dataset 
Cleaning the names 
Deleting last 3 columns (15, 16 and 17)
Dropping rows that detail mothly total (not specific to a dumpster)
Converting sports balls variable to an integer value 

```{r}
mr_trash_wheel = 
  readxl::read_excel(path = "./data/trash_wheel_data.xlsx") %>%
  janitor::clean_names() %>% 
  select(-x15, -x16, -x17) %>%
  drop_na(dumpster) %>% 
  mutate(sports_balls = as.integer(sports_balls))
```

Reading in the precipitation data for 2017
Adding a variable year = 2017 
Deleting rows without precipiatation data 

```{r}
ppt_2017 = 
  readxl::read_excel("./data/trash_wheel_data.xlsx", sheet = 6, range = "A2:B14") %>% 
  janitor::clean_names() %>%
  mutate(year = 2017)
```

Reading in the precipitation data for 2017
Adding a variable year = 2017 
Deleting rows without precipiatation data 

```{r}
ppt_2018 = 
  readxl::read_excel("./data/trash_wheel_data.xlsx", sheet = 5, range = "A2:B14") %>% 
  janitor::clean_names() %>%
  mutate(year = 2018)
```

Merging the datasets and convering month into a character variable 

```{r}
ppt_merged =
  bind_rows(ppt_2017, ppt_2018) %>%
  mutate(month = month.abb[month])
```


